---
title: "LCDB-symbology"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LCDB-symbology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(isaacr)
```

## Purpose

The goal of `lcdb_symbology` is to colour the symbology of the [Land Cover Database](https://lris.scinfo.org.nz/layer/104400-lcdb-v50-land-cover-database-version-50-mainland-new-zealand/) (LCDB) in a way that is consistent with the [Manaaki Whenua Landcare Research](https://www.landcareresearch.co.nz/) (MWLR)'s original symbology. MWLR provide a .lyr file for use with ArcGIS (or with QGIS via a plugin), but this repo provides a way to use the same symbology in R or ggplot2 or other non-GIS software.

## Usage

To load the LCDB symbology, use the `load_lcdb_symbology` function. This will return a data frame with the following columns: `Class`, `Colour`, and `Name`. The `Class` column contains the class code, the `Colour` column contains the colour in HEX format, and the `Name` column contains the name of the land cover class.

```{r setup}
# load data 
landcover_df <- load_lcdb_symbology()
```

## Table of colours

Here's a table of the colours used in the LCDB symbology. Extracted from XML in the original .lyr file. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(kableExtra)

# Display the table using kableExtra
landcover_df |> 
  mutate(
    ColourCell = cell_spec(Name, background = Colour, color = ifelse(Colour == "#000000", "white", "black"))
  ) %>%
  select(Class, ColourCell, Colour, Name) %>%
  kable("html", escape = FALSE, col.names = c("Class Code", "Colour", "HEX Code", "Landcover Class")) %>%
  kable_styling(full_width = FALSE, position = "left")
```

## Examples

Here's an example of how to use the LCDB symbology in ggplot2. Let's make a map of Christchurch. 

```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
library(ggplot2)
library(sf)

# for ggplot2, load named vector
lcdb_colours_vector <- load_lcdb_vector()

# Plot using ggplot2 with scale_fill_manual
ggplot(data = lcdb5_chch) + # lcdb5_chch available in the isaacr package, and it lazy-loads - so just refer to it
  geom_sf(aes(fill = as.factor(Class_2018)), color = NA) +
  scale_fill_manual(values = lcdb_colours_vector) +
  theme_void() +
  theme(legend.position = "none")
```
\
And another one of Blenheim

```{r fig.height=6, fig.width=8, message=FALSE, warning=FALSE}
# Plot using ggplot2 with scale_fill_manual
ggplot(data = lcdb5_blenheim) +
  geom_sf(aes(fill = as.factor(Class_2018)), color = NA) +
  scale_fill_manual(values = lcdb_colours_vector) +
  theme_void() +
  theme(legend.position = "none")
```

